Efficient Handling of String-Number Conversion

### 2019-11-21 draft.tex

Symbolic execution is a very popular technique that allows programmers to check the feasibility of a path in a given program, i.e., determining the value of the inputs under which the given path can be executed.
The path feasibility problem is usually solved by a reduction to the satisfiability of a formula. More precisely, program statements in the path are  translated to equivalent constraints in static single assignment (SSA) form and then solved by \emph{Satisfiability Modulo Theory (SMT)} solvers. The types of constraints needed depend on the types of program expressions to be analyzed. Therefore, SMT solvers need to support different combinations of theories so that they can handle a wide range of  types.

Among all data types, the \emph{string data type} is omnipresent in modern programming languages. Various security vulnerabilities such as injection and cross-site scripting attack are caused by malicious string values. Therefore, string constraint solving has received considerable attention in the constraint solving community. 
Operations such as \emph{equality constraints} (e.g. $x.y = y.x$), \emph{regular constraints} (e.g., $x \in (a.b)^*$), and \emph{integer constraints} (e.g., $|x|-|y|>3$), are widely supported by most state-of-the-art string constraint solvers such as, CVC4 \cite{?}, OSTRICH \cite{}, Sloth \cite{},  Trau+ \cite{abdulla2018trau,parosh2019chain,abdulla2017flatten}, Z3 \cite{?} and Z3Str3 \cite{}.

An important class of string operations is  the string-number conversions.  While string length operations are sometimes well supported, converting a string $x$ to an integer $n$ (e.g., using the operation $n = \sti{x}$) or turning an integer value $n$ into its string form $x$ (e.g., using the operation $x = \its{n}$)  suffer from extremely limited support by the state-of-the-art string constraint solvers. In fact, for many languages such as Java, string-number conversions are core operations, and much symbolic execution can not been done without supporting it. In many security vulnerable codes,  that receive input, tend to need to convert at least some of that input into numbers, which makes string-integer conversions critical to checking it. \yfc{some grammar error?} For example, the program fragment below is a variant of the Luhn test algorithm \cite{?} that is often used in credit card or personal ID validation.

>>>>

Symbolic execution is a very popular technique that allows programmers to check the feasibility of a path in a  program, i.e., determining the value of the inputs under which the given path can be executed.
The path feasibility problem is usually solved by a reduction to the satisfiability of a formula. More precisely, program statements in the path are  translated to equivalent constraints in static single assignment (SSA) form and then solved by \emph{Satisfiability Modulo Theory (SMT)} solvers. The types of constraints needed depend on the types of program expressions to be analyzed. Therefore, SMT solvers need to support different combinations of theories so that they can handle a wide range of  types.

Among all data types, the \emph{string data type} is omnipresent in modern programming languages. Various security vulnerabilities such as injection and cross-site scripting attack are caused by malicious string values. Therefore, string constraint solving has received considerable attention in the constraint solving community. 
Operations such as \emph{equality constraints} (e.g. $x.y = y.x$), \emph{regular constraints} (e.g., $x \in (a.b)^*$), and \emph{integer constraints} (e.g., $|x|-|y|>3$), are widely supported by most state-of-the-art string constraint solvers such as, CVC4 \cite{cvc4Tool}, OSTRICH \cite{chen2019decision}, Sloth \cite{sloth},  Trau+ \cite{abdulla2018trau,parosh2019chain,abdulla2017flatten}, Z3 \cite{z3} and Z3Str3 \cite{berzish2017z3str3}.

An important class of string operations is  the string-number conversions.  While string length operations are sometimes well supported, converting a string $x$ to an integer $n$ (e.g., using the operation $n = \sti{x}$) or turning an integer value $n$ into its string form $x$ (e.g., using the operation $x = \its{n}$)  suffer from extremely limited support by the state-of-the-art string constraint solvers.

For many languages such as Java, string-number conversions are core operations, and much symbolic execution can not been done without supporting it. In fact,  a code that receives input tends to need to convert at least some of that input into numbers,  which makes the handling of string-integer conversions of the utmost importance. For example, the program fragment below is a variant of the Luhn test algorithm \cite{?} that is often used in credit card or personal ID validation.
